---
title: "Food Rankings 2023"
author: "Amy Shuff"
output: word_document
date: '`r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Our 2023 Food Rankings analyze data from the 2022 Program Year (2021-2022 data).


``` {r, echo=FALSE, warning=FALSE, results='hide', message = FALSE}
# March 2023 Update
# Last year the file "Sample Dataset Template_PY21 SNP Meals_ChildrenAtRisk_11152021" outlined how the TDA data was manipulated in Excel before being imported into R. The goal with this update is to do as much work as possible in R.

library(usethis)
library(tidyverse)
library(janitor)
library(reshape2)
library(stringr)
library(here)
library(readxl)
```

# Data Sources

## Harmony and Kipp Districts

Explain here why we have to combine harmony and kipp districts. Or maybe this should go someplace else?

```{r, echo=FALSE, warning=FALSE, results='hide', message = FALSE}
harmony <- c(101858, 101862, 71806, 15828, 161807, 101846, 227816) #use to combine Harmony districts
kipp_data <- c(227820, 57837, 101813, 15826) #use to combine KIPP districts

```

## Texas Education Agency (TEA)

Total Enrollment Counts in Student Program and Special Populations Reports, PEIMS Data 2021-2022

TEA data was obtained via Public Records Request. From this data we use the student enrollment counts and economically disadvantaged counts.

Student enrollment for small campuses are masked. In the case of the reported number being <10, we use 10.
This means our district student counts are a slight overestimate.

Economically disadvantaged counts are sometimes unavailable, so in that case they are assumed zero.

Percent economically disadvantaged for each district is calculated from the campus counts.

``` {r, echo=FALSE, warning=FALSE, results='hide', message = FALSE}

#This is something we request for the school rankings.

tea <- read_xlsx("pru_7128_StudPgmStateCampus22state.xlsx", skip = 6)
  
#Small enrollment numbers are masked with numbers like <10, so here we'll remove all the "<" characters.
#Enrollment for small campuses will be an over estimate.
tea$"ALL ENROLLMENT" =  gsub("<", "", tea$"ALL ENROLLMENT")

tea <- tea %>% 
  mutate(ENROLLMENT = as.numeric(tea$"ALL ENROLLMENT"),
         #Economically disadvantaged numbers are masked with "-999"
         eco.dis = ifelse(tea$"ECONOMICALLY DISADVANTAGED" >= 0, tea$"ECONOMICALLY DISADVANTAGED", NA)) %>%
  rename(District.Number = "DISTRICT NUMBER") %>% 
  group_by(District.Number) %>% 
  summarise(tea.all.stud = sum(ENROLLMENT),
            tea.eco.dis = sum(eco.dis, na.rm = T)) %>% 
  mutate(eco.dis.pct = (tea.eco.dis/tea.all.stud)*100)
```



## Texas Department of Agriculture (TDA)

Although we obtained our data from a public information request, it is also available on the Food and Nutrition Meals Served Dashboard, 2021-2022: https://data.texas.gov/stories/s/2021-2022-TDA-Food-and-Nutrition-Meals-Served-Dash/93tt-ffn6

### School Nutrition Program (SNP)

SNP Data includes meal counts from the National School Lunch Program (NSLP) and the School Breakfast Program (SBP).

Credit for participating in the Community Eligible Provision (CEP) was given to the entire district if they had any participating location.


``` {r, echo=FALSE, warning=FALSE, results='hide', message = FALSE}

# This year TDA School Nutrition Program Data is from our public records request.
# It comes in an excel sheet with multiple tabs. We'll import each of the tabs, select the variables we want to keep, then join them together.
# This will probably look different next year. First check the Texas Open Data Portal for data before submitting records request.
# Notes from last year say that only the universal breakfast waiver information wasn't available without PIR
# Use this link to look for next year's data: https://data.texas.gov/stories/s/e2dm-5r4v

# Contacts and Participation has CEP data
snp.contacts <- read_xlsx(here("WF Attachment 180587 PIR 23-194_PY22 SNP Meal Detail_01312023.xlsx"), sheet = 2) %>%
  mutate(CEID = as.numeric(CEID),
         SiteID = as.numeric(SiteID),
         ESC.Region = as.numeric(ESC)) %>%
  group_by(CEID, SiteID) %>%
  rename(County = CECounty,
        District = CountyDistrictCode) %>% 
  select(CEID, SiteID, CEName, District, ESC.Region,
         SchoolBreakfastProgram, NationalSchoolLunchProgram, AfterschoolCareProgram,
         SevereNeedBreakfast, UniversalFreeBreakfast, AreaEligibleSnack, CEP, Provision2)

#Meal Reimbursement Claims has the number of meals served
#this tab has claims each month for each site
#the work here represents what was done in excel last year
snp.meal.reimb <- read_xlsx(here("WF Attachment 180587 PIR 23-194_PY22 SNP Meal Detail_01312023.xlsx"), sheet = 3) %>%
  mutate(CEID = as.numeric(CEID),
         SiteID = as.numeric(SiteID),
         ESC.Region = as.numeric(ESC),
         ClaimDate = as.character(ClaimDate),
         
         #tda notes say that they consider october the typical snapshot for the year
         #this is what they did last year
         # EnrollmentQty = ifelse(ClaimDate == "2021-10-01", EnrollmentQty, 0),
         # FreeEligQty = ifelse(ClaimDate == "2021-10-01", FreeEligQty, 0),
         # RedcEligQty = ifelse(ClaimDate == "2021-10-01", RedcEligQty, 0),
         # PaidEligQty = ifelse(ClaimDate == "2021-10-01", PaidEligQty, 0)
         #only 455 sites submitted claims for October, though
         
         ) %>% 
  group_by(CEID, SiteID) %>%
  summarise(
        CEName = first(CEName),
        County = first(CECounty),
        District = first(CountyDistrictCode),
        ESC.Region = first(ESC.Region),
        
        # this sum would find the october number, if we were using that method
        # Enrollment_TDA = sum(EnrollmentQty, na.rm = T),
        # Eligible.Free = sum(FreeEligQty, na.rm = T),
        # Eligible.Reduced = sum(RedcEligQty, na.rm = T),
        
        # sums here represent site total for year
        
        SBP.Days.Served = sum(BreakfastDays, na.rm = T),  
        NSLP.Days.Served = sum(LunchDays, na.rm = T),
        NSLP.Snack.Days.Served = sum(SnackDays, na.rm=T),
        # SBP.Free = sum(BreakfastServedFree, na.rm=T),    
        # SBP.Reduced = sum(BreakfastServedRedc, na.rm=T),   
        # NSLP.Free = sum(LunchServedFree, na.rm=T),
        # NSLP.Reduced = sum(LunchServedRedc, na.rm=T),
        
        
        #Because for this particular year everyone could get free meals, I'm assuming those eligible for paid meals still didn't pay.
        #so I'm going to look at all meals, regardless of eligibility
        NSLP.Total = sum(LunchTotal),
        SBP.Total = sum(BreakfastTotal)
        )

snp <- full_join(snp.contacts, snp.meal.reimb, by = c("CEID", "SiteID", "CEName", "District", "ESC.Region"))


#I don't think we use the cacfp contacts tab for anything
#cacfp.contacts <- read_xlsx(here("WF Attachment 180587 PIR 23-194_PY22 SNP Meal Detail_01312023.xlsx"), sheet = 4)


#Not sure if we need this. These schools didn't serve universal breakfast even though they received money for free lunches.
#District granted waiver from Senate Bill 376: Lists of CEs and sites approved to waive the Universal Breakfast requirement for PY22.
#Before importing this data to R, manually check and add CEID and Site ID if not already in sheet
#ub.waiver <- read_xlsx(here("WF Attachment 180587 PIR 23-194_PY22 SNP Meal Detail_01312023.xlsx"), sheet = 5) %>%
#  rename(SiteID = "SITE ID", CEName = "CENAME") %>%
#  mutate(CEID = as.numeric(CEID),
#         SiteID = as.numeric(SiteID),
#         UniversalBreakfastWaiver = "Y") %>% 
#  select(CEID, SiteID, CEName, UniversalBreakfastWaiver)
#snp.full <- full_join(snp, ub.waiver, by = c("CEID", "SiteID", "CEName"))
```




### Seamless Summer Option (SSO)

SSO data was obtained from the TDA F&N Meals Served Dashboard.

2021-2022 Monthly SNP & SSO School Meal Count:
https://data.texas.gov/dataset/2021-2022-Monthly-School-Meal-Count/dyrd-amq2

Note from the data dashboard:
"In March 2020, USDA began allowing flexibility in nutrition assistance program policies to support continued meal access during the coronavirus pandemic (COVID-19). This included offering School Nutrition Program sponsors the option to provide meals through Seamless Summer Option (SSO) in program year 2021-2022." 

The SSO is what schools used to offer free meals to all students regardless of income. Because of this, we are choosing to look at total number of meals served instead of only the free and reduced as we've done in years past. This option will not be available next year because the program has ended. 


``` {r, echo=FALSE, warning=FALSE, results='hide', message = FALSE}

#This waiver has ended, so disregard this part next year.

sso <- read.csv("2021-2022_School_Meal_Count_-_TDA_F_N_Dashboard.csv") %>% 
  filter(ProgramYear == "2021-2022",
         Program == "SEAMLESS_SUMMER_OPTION_MEALS") %>% 
  group_by(CEID, SiteID) %>%
  summarise(
        CEName = first(CEName),
        County = first(CECounty),
        District = first(CountyDistrictCode),
        ESC.Region = first(ESC),
        
        #sums here represent site total for year
        SSO.Breakfast = sum(BreakfastTotal, na.rm = T),  
        SSO.Lunch = sum(LunchTotal, na.rm = T),
        SSO.Breakfast.Days = sum(BreakfastDays, na.rm = T),  
        SSO.Lunch.Days = sum(LunchDays, na.rm = T),
        SSO.Snack.Days = sum(AMSnackDays + PMSnackDays, na.rm=T),
        )
```

### Child and Adult Care Food Program (CACFP)

Child and Adult Care Food Programs (CACFP) – At Risk– Meal Reimbursement – Program Year 2021-2022:
https://data.texas.gov/dataset/Child-and-Adult-Care-Food-Programs-CACFP-At-Risk-M/e4wr-4i5j

CACFP data captures meals served in At-Risk afterschool care centers.

``` {r, echo=FALSE, warning=FALSE, results='hide', message = FALSE}
#Next year, use this link: https://data.texas.gov/stories/s/iekx-7mdi
#and scroll down to CACFP At-Risk Child Care Claims


cacfp <- read.csv(here("Child_and_Adult_Care_Food_Programs__CACFP____At_Risk__Meal_Reimbursement___Program_Year_2021-2022.csv")) %>%
  mutate(CEID = as.numeric(CEID),
         SiteID = as.numeric(SiteID),
         ESC.Region = as.numeric(ESC)) %>%
  group_by(CEID, SiteID) %>%
  summarise(CEName = first(CEName),
            County = first(CECounty),
            District = first(CountyDistrictCode),
            ESC.Region = first(ESC.Region),        
            CACFP.at.Risk.Supper.Days.Served = sum(SupperDays, na.rm=T),
            CACFP.at.Risk.Snack.Days.Served = sum(SnackDays, na.rm=T))
```




# Aggregate school-level data up to district level

``` {r, echo=FALSE, warning=FALSE, results='hide', message = FALSE}
district <- full_join(snp, cacfp, by=c("CEID", "SiteID", "CEName", "District", "ESC.Region", "County")) %>% 
  full_join(sso, by=c("CEID", "SiteID", "CEName", "District", "ESC.Region", "County"))

district$District = gsub("-", "", district$District)

district <- district %>% 
  mutate(District = as.numeric(District)) %>% 
  mutate(
          CEP01 = ifelse(CEP=="Y", 1, 0),
          District.Number = ifelse(District %in% harmony, 227816,
                                   ifelse(District %in% kipp_data, 227820, District))
          ) %>% 
  group_by(District.Number) %>% 
  summarise(
          CEName = first(CEName),
          County = first(na.omit(County)),
          ESC.Region = first(ESC.Region),
          #Eligible.Free = sum(Eligible.Free, na.rm = T),
          #Enrollment_TDA = sum(Enrollment_TDA, na.rm = T),
          #Eligible.Reduced = sum(Eligible.Reduced, na.rm = T),
          NSLP.Days.Served = max(NSLP.Days.Served, 0, na.rm = T),
          #NSLP.Free = sum(NSLP.Free, na.rm=T),
          #NSLP.Reduced = sum(NSLP.Reduced, na.rm=T),
          SBP.Days.Served = max(SBP.Days.Served, 0, na.rm = T),
          #SBP.Free = sum(SBP.Free, na.rm=T),
          #SBP.Reduced = sum(SBP.Reduced, na.rm=T),
          CACFP.at.Risk.Supper.Days.Served = sum(CACFP.at.Risk.Supper.Days.Served, na.rm=T),
          CACFP.at.Risk.Snack.Days.Served = sum(CACFP.at.Risk.Snack.Days.Served, na.rm=T),
          NSLP.Snack.Days.Served = sum(NSLP.Snack.Days.Served, na.rm=T),
          CEP = sum(CEP01, na.rm=T),
          
          SBP.Total = sum(SBP.Total, na.rm=T),
          NSLP.Total = sum(NSLP.Total, na.rm=T),

          #Next year we don't need SSO numbers
          SSO.Breakfast = sum(SSO.Breakfast, na.rm = T),  
          SSO.Lunch = sum(SSO.Lunch, na.rm = T),
          SSO.Breakfast.Days = max(SSO.Breakfast.Days, 0, na.rm = T),
          SSO.Lunch.Days = max(SSO.Lunch.Days, 0, na.rm = T),
          SSO.Snack.Days = max(SSO.Snack.Days, 0, na.rm = T)    
          ) %>%
  
  left_join(., tea, by="District.Number") %>% 
  mutate(
          CEName = ifelse(District.Number==227820, "KIPP SCHOOLS", CEName),
          #Eligible.Free and Eligible.Reduced are from the month with the greatest claims from schools
          #next year use october numbers
          #tda.frl.total = Eligible.Free + Eligible.Reduced,
         # NSLP.Days.Served = ifelse(is.na(NSLP.Days.Served), 0, NSLP.Days.Served),
         # SBP.Days.Served = ifelse(is.na(SBP.Days.Served), 0, SBP.Days.Served),
          
          #don't need SSO next year, and instead of total, use free and reduced instead
          Breakfast.Served = SSO.Breakfast + SBP.Total,
          Breakfast.Days = SSO.Breakfast.Days + SBP.Days.Served,
          Lunch.Served = SSO.Lunch + NSLP.Total,
          Lunch.Days = SSO.Lunch.Days + NSLP.Days.Served,
          Snack = SSO.Snack.Days + NSLP.Snack.Days.Served + CACFP.at.Risk.Snack.Days.Served,
          
          #ADP is Average Daily Participation: The average number of students that ate on days meals were offered.         
          Breakfast.ADP = ifelse(Breakfast.Days==0, 0, Breakfast.Served/Breakfast.Days),
          Lunch.ADP = ifelse(Lunch.Days==0, 0, (Lunch.Served/Lunch.Days)),
          
          #Participation Rates are the percent of students that ate
          Pct.Breakfast.Participation = (Breakfast.ADP/tea.all.stud)*100,
          Pct.Lunch.Participation = (Lunch.ADP/tea.all.stud)*100,
          cacfp_supper01 = ifelse(CACFP.at.Risk.Supper.Days.Served>0, 100, 0),
          snack_anyafter =  ifelse(Snack>0, 100, 0)
                )
```


# Overall Rankings

``` {r, echo=FALSE, warning=FALSE, results='hide', message = FALSE}
district_rankings <- district %>% 
  #remove small districts (less than 10,000 students) and those that aren't at least 60% economically disadvantaged
  filter(tea.all.stud>=10000 & eco.dis.pct >= 60) %>% 
  
  #This is where we find our overall score for food rankings
  mutate(district_sum = ((Pct.Lunch.Participation*0.25)+(Pct.Breakfast.Participation*0.5)+(cacfp_supper01*0.1)+(snack_anyafter*0.15))) %>% 
  
  arrange(-district_sum) %>% 
  mutate(Rank = row_number()) %>% 
  select(CEName, Rank, County, ESC.Region, tea.all.stud, eco.dis.pct, district_sum, Pct.Lunch.Participation, 
         Pct.Breakfast.Participation, cacfp_supper01, snack_anyafter, CEP) %>%
  mutate(cacfp_supper01 = ifelse(cacfp_supper01==100, "Yes", "No"),
         snack_anyafter = ifelse(snack_anyafter==100, "Yes", "No"),
         CEP = ifelse(CEP>0, "Yes", "No")) %>% 
  mutate_if(is.numeric, round, digits = 1) %>% 
  reshape::rename(c(CEName = "District Name",
                    ESC.Region = "ESC Region",
                    tea.all.stud = "Total Enrollment",
                    eco.dis.pct = "% Economically Disadvantaged",
                    district_sum = "Overall Score",
                    Pct.Lunch.Participation = "% Lunch Participation",
                    Pct.Breakfast.Participation = "% Breakfast Participation",
                    cacfp_supper01 = "CACFP Supper",
                    snack_anyafter = "Afterschool Snack")) 

write_csv(district_rankings, here("foodrankings_PY22_10k.csv"))
```


# Large districts, 50k or more
Creates same table as before but with 50K or more students
``` {r, echo=FALSE, warning=FALSE, results='hide', message = FALSE}
district_50k <- district %>% 
  filter(tea.all.stud>=50000 & eco.dis.pct >= 60) %>% 
  mutate(district_sum = ((Pct.Lunch.Participation*0.25)+(Pct.Breakfast.Participation*0.5)+(cacfp_supper01*0.1)+(snack_anyafter*0.15))) %>% 
  arrange(-district_sum) %>% 
  mutate(Rank = row_number()) %>% 
  select(CEName, Rank, County, ESC.Region, tea.all.stud, eco.dis.pct, district_sum, Pct.Lunch.Participation, 
         Pct.Breakfast.Participation, cacfp_supper01, snack_anyafter, CEP) %>%
  mutate(cacfp_supper01 = ifelse(cacfp_supper01==100, "Yes", "No"),
         snack_anyafter = ifelse(snack_anyafter==100, "Yes", "No"),
         CEP = ifelse(CEP>0, "Yes", "No")) %>% 
  mutate_if(is.numeric, round, digits = 1) %>% 
  reshape::rename(c(CEName = "District Name",
                    ESC.Region = "ESC Region",
                    tea.all.stud = "Total Enrollment",
                    eco.dis.pct = "% Economically Disadvantaged",
                    district_sum = "Overall Score",
                    Pct.Lunch.Participation = "% Lunch Participation",
                    Pct.Breakfast.Participation = "% Breakfast Participation",
                    cacfp_supper01 = "CACFP Supper",
                    snack_anyafter = "Afterschool Snack")) 

write_csv(district_50k, here("foodrankings_PY22_50k.csv"))
```

# Middle-income districts (60-70% Eco Dis)
takes out high poverty schools and focuses on those which often face more challenges in implementing school food programs to scale
``` {r, echo=FALSE, warning=FALSE, results='hide', message = FALSE}
district_middleincome <- district %>% 
  filter(tea.all.stud>=10000 & eco.dis.pct >= 60 & eco.dis.pct <=70) %>% 
  mutate(district_sum = ((Pct.Lunch.Participation*0.25)+(Pct.Breakfast.Participation*0.5)+(cacfp_supper01*0.1)+(snack_anyafter*0.15))) %>% 
  arrange(-district_sum) %>% 
  mutate(Rank = row_number()) %>% 
  select(CEName, Rank, County, ESC.Region, tea.all.stud, eco.dis.pct, district_sum, Pct.Lunch.Participation, 
         Pct.Breakfast.Participation, cacfp_supper01, snack_anyafter, CEP) %>%
  mutate(cacfp_supper01 = ifelse(cacfp_supper01==100, "Yes", "No"),
         snack_anyafter = ifelse(snack_anyafter==100, "Yes", "No"),
         CEP = ifelse(CEP>0, "Yes", "No")) %>% 
  mutate_if(is.numeric, round, digits = 1) %>% 
  reshape::rename(c(CEName = "District Name",
                    ESC.Region = "ESC Region",
                    tea.all.stud = "Total Enrollment",
                    eco.dis.pct = "% Economically Disadvantaged",
                    district_sum = "Overall Score",
                    Pct.Lunch.Participation = "% Lunch Participation",
                    Pct.Breakfast.Participation = "% Breakfast Participation",
                    cacfp_supper01 = "CACFP Supper",
                    snack_anyafter = "Afterschool Snack")) 

write_csv(district_middleincome, here("foodrankings_PY22_middleincome.csv"))
```